---
title: "My Sample R Codes"
output: html_document
Author: Sedi Abbasnasab
date: "2022-10-01"
---

# LOAD DATA 

```{r}
library(datasets)
```

# SUMMARIZE DATA

```{r}
head(iris)
summary(iris)
plot(iris)
```

# CLEAN UP
# Clear packages

```{r}
detach("package:datasets", unload = TRUE)
```

# Clear plots

```{r}
dev.off()
```

# Clear console

Ctrl+L

# LOAD PACKAGES
# Install packages

```{r}
install.packages("pacman")
library(pacman)
```

# These are the packages I load every time.

```{r}
pacman::p_load(pacman, dplyr, GGally, ggplot2, ggthemes, ggvis, httr, lubridate, plotly, rio, rmarkdown, shiny, stringr, tidyr)
```

# LOAD BASE PACKAGES

```{r}
library(datasets)
```

# LETS LOOK AT THE iris DATASET

```{r}
head(iris)
```

# CREATE PLOTS
# Categorical variable

```{r}
plot(iris$Species)
```

# Quantitative variable

```{r}
plot(iris$Petal.Length)
```

# Cat + Quant

```{r}
plot(iris$Species, iris$Petal.Width)
```

# Quant pair

```{r}
plot(iris$Petal.Length, iris$Petal.Width)
```

# Enttire data frame

```{r}
plot(iris)
```

# Plot with options

```{r}
plot(iris$Petal.Length, iris$Petal.Width,
     col = "#cc0000", pch = 19, 
     main = "Iris: Petal Lenght vs. Petal Width",
     xlab = "Petal Lenght",
     ylab = "Petal Width")
```

# Plot formulas with plot()

```{r}
plot(cos, 0, 2*pi)
plot(exp, 1, 5)
plot(dnorm, -3, +3)
```

# Formula plot with options

```{r}
plot(dnorm, -3, +3,
     col = "#cc0000",
     Iwd = 5,
     main = "Standard Normal Distribution",
     xlab = "z-scores",
     ylab = "Density")
```

# Let's look at mtcars dataset

```{r}
library(datasets)
?mtcars
head(mtcars)
```

# Bar charts

```{r}
cylinders <- table(mtcars$cyl) #Create table
barplot(cylinders) # Bar chart
plot(cylinders) # Default X-Y plot (lines)
```

# Get back to iris dataset

```{r}
library(datasets)
?iris
head(iris)
```

# Basic histograms

```{r}
hist(iris$Sepal.Length)
hist(iris$Sepal.Width)
hist(iris$Petal.Length)
hist(iris$Petal.Width)
```

# Histogram for each species using options

```{r}
hist(iris$Petal.Width [iris$Species == "setosa"],
     xlim = c(0, 3),
     breaks = 9,
     main = "Petal Width for Setosa",
     xlab = "",
     col = "red")

hist(iris$Petal.Width [iris$Species == "versicolor"],
     xlim = c(0, 3),
     breaks = 9,
     main = "Petal Width for Versicolor",
     xlab = "",
     col = "purple")

hist(iris$Petal.Width [iris$Species == "virginica"],
     xlim = c(0, 3),
     breaks = 9,
     main = "Petal Width for Virginica",
     xlab = "",
     col = "blue")
```

# Restore graphic parameter

```{r}
par(mfrow=c(1, 1))
```

# Scatterplot using mtcars dataset

```{r}
library(datasets)
?mtcars
head(mtcars)
```

# Good to first check univariate distributions

``` {r}
hist(mtcars$wt)
hist(mtcars$mpg)
```

# Basic X-Y plot for two quantitative variables

```{r}
plot(mtcars$wt, mtcars$mpg)
```

# I have now added some options to the above plot

```{r}
plot(mtcars$wt, mtcars$mpg,
     pch = 19,
     cex = 1.5,
     col = "#cc0000",
     main = "MPG as a Function of Weight of Cars",
     xlab = "Weight (in 1000 pounds)",
     ylab = "MPG")
```

# OVERLAYING PLOTS
# I am going to use lynx dataset for demonstrating overlaying plots

```{r}
library(datasets)
?lynx
head(lynx)
```

# Histogram 
# Default

```{r}
hist(lynx)
```

# I am going to add some options to the above histogram and then add a normal distribution, add two kernel density estimators, finally add a rug plot

```{r}
hist(lynx,
     breaks = 14,
     freq = FALSE,
     col = "thistle1",
     main = paste("Histogram of Annual Canadian Lynx",
                  "Trappings, 1821-1934"),
     xlab = "Number of Lynx Trapped")
curve(dnorm(x, mean = mean(lynx), sd = sd(lynx)),
      col = "thistle4",
      lwd = 2,
      add = TRUE)
lines(density(lynx), col = "blue", lwd = 2)
lines(density(lynx, adjust = 3), col = "purple", lwd = 2)
rug(lynx, lwd = 2, col = "gray")
```

# summary

```{r}
library(datasets)
head(iris)
summary(iris$Species) # categorical variable
summary(iris$Sepal.Length) # quantitative variable
summary(iris) # Entire data frame
```

# describe
# Download the psych package first

```{r}
p_load(psych)
```
```{r}
describe(iris$Sepal.Length) # one quantitative variable
describe(iris) # entire data frame
```

# Selecting cases

```{r}
hist(iris$Petal.Length)
summary(iris$Petal.Length)
summary(iris$Species)
```

# Select by category

# Versicolor

```{r}
hist(iris$Petal.Length[iris$Species == "versicolor"],
     main = "Petal Lenght: Versicolor")
```

# Virginica

```{r}
hist(iris$Petal.Length [iris$Species == "virginica"],
     main = "Petal Lenght: Virginica")
```

# Setosa

```{r}
hist(iris$Petal.Length [iris$Species == "setosa"],
     main = "Petal Lenght: Setosa")
```

# Select by value

# Short petals only (all Setosa)

```{r}
hist(iris$Petal.Length [iris$Petal.Length < 2],
     main = "Petal Lenght < 2")
```

# Multiple selectors

# Short Virginica petals only

```{r}
hist(iris$Petal.Length [iris$Species == "virginica" &
iris$Petal.Length < 5.5],
main = "Petal Length: Short Virginica")
```
# Create subsample

```{r}
i.setosa <- iris[iris$Species == "setosa", ]
```

# Explore subsample

```{r}
head(i.setosa)
summary(i.setosa$Petal.Length)
hist(i.setosa$Petal.Length)
```

# Data types

# Numeric

```{r}
n1 <- 15 # Double precision by default
n1
typeof(n1)
```
```{r}
n2 <- 1.5
n2
typeof(n2)
```
# Character

```{r}
c1 <- "c"
c1
typeof(c1)
```
```{r}
c2<- "a string of text"
c2
typeof(c2)
```

# Logical

```{r}
l1 <- TRUE
l1
typeof(l1)
```
```{r}
l2 <- FALSE
l2
typeof(l2)
```

# Data structures

# Vector

```{r}
v1 <- c(1, 2, 3, 4, 5)
v1
is.vector(v1)
```
```{r}
v2 <- c("a", "b", "c", "d")
v2
is.vector(v2)
```
```{r}
v3 <- c(TRUE, TRUE, FALSE, FALSE, TRUE)
v3
is.vector(v3)
```

# Matrix

```{r}
m1 <- matrix(c(T, T, F, F, T, F), nrow = 2)
m1
```
```{r}
m2 <- matrix(c("a", "b", "c", "d"),
             nrow = 2,
             byrow = T)
m2
```

# Array

```{r}
a1 <- array(c(1:24), c(4, 3, 2))
a1
```

# Data frame

```{r}
vNumeric <- c(1, 2, 3)
vCharacter <- c("a", "b", "c")
vLogical <- c(T, F, T)
```
```{r}
dfa <- cbind(vNumeric, vCharacter, vLogical)
dfa <- as.data.frame(cbind(vNumeric, vCharacter, vLogical))
dfa
```

# List
```{r}
o1 <- c(1, 2, 3)
o2 <- c("a", "b", "c", "d")
o3 <- c(T, F, T, T, F)

list1 <- list(o1, o2, o3)
list1

list2 <- list(o1, o2, o3, list1)
list2
```

# Automatic coercion

# Goes to "least restrictive" data type

```{r}
(coerce1 <- c(1, "b", TRUE))
typeof(coerce1)
```

# Coerce numeric to integer

```{r}
(coerce2 <- 5)
typeof(coerce2)

(coerce3 <- as.integer(5))
typeof(coerce3)
```

# Coerce matrix to data frame

```{r}
(coerce4 <- matrix(1:9, nrow= 3))
is.matrix(coerce4)

(coerce5 <- as.data.frame(matrix(1:9, nrow= 3)))
is.data.frame(coerce5)
```

# Create data

```{r}
(x1 <- 1:3)
(y <- 1:9)
# Combine variables
(df1 <- cbind.data.frame(x1, y))
typeof(df1$x1)
```

# As.Factor

```{r}
(x2 <- as.factor(c(1:3)))
(df2 <- cbind.data.frame(x2, y))
typeof(df2$x2)
str(df2)
```

# Define existing variable as factor

```{r}
x3 <- c(1:3)
df3 <- cbind.data.frame(x3, y)
(df3$x3 <- factor(df3$x3, levels = c(1, 2, 3)))
typeof(df3$x3)
str(df3)
```

# Lables for factor

```{r}
x4 <- c(1:3)
df4 <- cbind.data.frame(x4, y)
df4$x4 <- factor(df4$x4,
                 levels = c(1, 2, 3),
                 labels = c("macOS", "Windows", "Linux"))
df4
typeof(df4$x4)
str(df4)
```

# Ordered factors and labels

```{r}

